---
title: "Exploratory Data Analysis"
output:
  pdf_document: default
  html_notebook: default
---


### Base Plotting Functions

- `plot`: scatterplot or other kind of plot depending on the class of the object being plotted
- `lines`: Adds lines to an existing plot
- `points`: Adds points to an existing plot
- `text`: Add text labels to a plot using `x,y` coordinates
- `title`: Adds annotations to axis, titles, subtitles or outer margin
- `mtext`: Adds text to the margins
- `axis`: Adds axis ticks or labels


### Base Plot with an Annotation


```{r}
library(datasets)
with(airquality, plot(Wind,Ozone))
title(main="Ozone and Wind in New York City")
```


```{r}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City"))
with(subset(airquality, Month==5), points(Wind, Ozone, col = "blue"))
```

```{r}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", type="n"))
with(subset(airquality, Month==5), points(Wind, Ozone, col = "blue"))
with(subset(airquality, Month!=5), points(Wind, Ozone, col = "red"))
legend("topright", pch=1, col = c("blue", "red"), legend = c("May", "Other Months"))
```


### Base Plot with Regression Line

```{r}
with(airquality, plot(Wind, Ozone, main = "Ozone and Wind in New York City", pch = 20))
model <- lm(Ozone ~ Wind, airquality)
abline(model, lwd = 2)
```


### Multiple Base Plots

```{r}
par(mfrow = c(1,2))
with(airquality, {
    plot(Wind, Ozone, main = "Ozone and Wind")
    plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
})
```

```{r}
par(mfrow = c(1,3), mar=c(4,4,2,1), oma = c(0,0,2,0))
with(airquality, {
    plot(Wind, Ozone, main = "Ozone and Wind")
    plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
    plot(Temp, Ozone, main = "Ozone and Temperature")
    mtext("Ozone and Weather in New York City", outer = TRUE)
})
```

## Base Plotting Demonstrations

```{r}
x <- rnorm(100)
hist(x)
y <- rnorm(100)
plot(x,y, pch = 6)
fit <- lm(y ~ x)
abline(fit, lwd = 3, col = "red")

```
 
 
 
 
```{r}
x <- rnorm(100)
y <- x + rnorm(100)
g <- gl(2,50)
g <- gl(2, 50, labels = c("Male", "Female"))
str(g)
plot(x, y, type = "n") # Make the plot but do not put any data in there
points(x[g == "Male"], y[g == "Male"], col = "green")
points(x[g == "Female"], y[g == "Female"], col = "red")
```


## Graphic Devices on R

### How to create a Plot

```{r}
library(datasets)
with(faithful, plot(eruptions, waiting))
tile(main = "Old Faithful Geyser Data")
```

Second approach for file devices:


```{r}
pdf(file="example.pdf")
with(faithful, plot(eruptions,waiting))
title(main="Old Faithful Geysers data")
dev.off()
```

### Graphics File devies


Vector formats:

* `pdf`

* `svg`: XML-based scalable vector graphics.

* `win.metafile`: Windows Metafile (only for Windows)

* `postscript`

### Copying Plots

* `dev.copy`

* `dev.copy2pdf`: copy a plot to a PDF file

```{r}
library(datasets)
with(faithful, plot(eruptions, waiting))
title(main = "Old Faithful Geyser Data")
dev.copy(png, file = "geyserplot.png")
dev.off()
```




## Week 1 Quiz

1. Which of the following is a principle of analytic graphics? X Don't plot more than two variables at a time

2. What is the role of exploratory graphs in data analysis? Personal Understanding

3. Which of the following is true about the base plotting system? Plots are created and annotated with separate functions

4. Which of the following is an example of a valid graphics device in R? PNG 

5. Which of the following is an example of a vector graphics device in R? SVG

6. Bitmapped file formats can be most useful for Scatterplots with many many points

7. Which of the following functions is typically used to add elements to a plot in the base graphics system? X plot()

8. Which function opens the screen graphics device for the Mac? Quartz()

9. What does the 'pch' option to par() control? the plotting symbol/character in the base graphics system

10. If I want to save a plot to a PDF file, which of the following is a correct way of doing that? Construct the plot on the screen device and then copy it to a PDF file with dev.copy2pdf()
