---
title: "R Notebook"
output: html_notebook
---

# Week 2 Quiz

## Question 1

```{r}
library(lattice)
g <- bwplot(hwy ~ displ, mpg)
#summary(mpg)
class(g)
```

## Question 2

```{r}
library(nlme)
library(lattice)
xyplot(weight ~ Time | Diet, BodyWeight)
```

Weight vs time for each diet.

## Question 3

```{r}
```


llines? CORRECT
lpoints?

annotate panels: 

text:The text() function is used to annotate plots created with the base plotting system. wrong!

## Question 4

```{r}
library(lattice)
library(datasets)
data(airquality)
p <- xyplot(Ozone ~ Wind | factor(Month), data = airquality)
p
```

## Question 5

splom WRONG!
print.trellis() wrong
trellis.par.set() CORRECT

## Question 6

Grammar of Graphics

## Question 7

I am interested in examining how the relationship between ozone and wind speed varies across each month. What would be the appropriate code to visualize that using ggplot2? 

-- Use the one with transform ---

```{r}
library(datasets)
data(airquality)
airquality = transform(airquality, Month = factor(Month))
qplot(Wind, Ozone, data = airquality, facets = . ~ Month)
```


vs

```{r}
qplot(Wind, Ozone, data = airquality, facets = . ~ factor(Month))
```

## Question 8

plloting object

## Question 9

```{r}
library(ggplot2)
library(ggplot2movies)
g <- ggplot(movies, aes(votes, rating))
g+geom_point()
```

what type of layer?


## Question 10

```{r}
qplot(votes, rating, data = movies)
qplot(votes, rating, data = movies) + geom_smooth()
```

