---
title: "Exploratory Data Analysis - Week 2"
output:
  html_notebook: default
  pdf_document: default
---

```{r "setup", include=FALSE}
require("knitr")
library(ggplot2)
library(lattice)
library('lubridate')
library('dplyr')
```


```{r}
plot(cars)
```


## Week 2

### The Lattice Plotting system

Implemented with the following packages:

- `lattice`
- `grid`

#### Main functions

- xyplot
- bwplot
- histogram
- stripplot
- dotplot
- splom
- levelplot / counterplot


#### xyplot

The function is used in the following way

`xyplot(y ~ x | f * g, data)`

- y is the y-axis variable
- x is the x-axis variable
- f and g, condition variables
- data is the data frame where the variables should be extracted from 

```{r}
library(lattice)
xyplot(Temp ~ Month, data = airquality)
```

Now converting the month variable into a factor

```{r}
library(datasets)
library(lattice)
airquality <- transform(airquality, Month = factor(Month))
xyplot( Ozone ~ Wind | Month, data = airquality, layout = c(5,1))
```

### Lattice Panel Functions

```{r}
set.seed(10)
x <- rnorm(100)
f <- rep(0:1, each=50)
y <- x + f - f * x + rnorm(100, sd = 0.5)
f <-factor(f, labels = c("Group 1", "Group 2"))
xyplot (y ~ x | f, layout = c(2,1))
```

#### Custom Panel function

```{r}
xyplot (y ~ x | f, panel = function(x,y, ...) {
  panel.xyplot(x, y, ...) ## First call the default panel function for xyplot
  #panel.abline( h = median(y), lty = 2) ## Add a horizontal line at the median
  panel.lmline(x, y, col = 2)
})
```


### Working with colors (swirl)

We use the library grDevices:

```{r}
library(grDevices)

```

We can use the function colorRamp to return another function which will make a specific number of gradients.

`colorRampPalette` takes a palette of colros and returns a function.

```{r}
p1 <- colorRampPalette(c("red", "blue"))
p1(2) ## returns a vector of two colors
p1(6) ## returns a vector of six colors
```

The brewer package contains different color palettes, from colors center to the white, to divergent to different. (colorblind?)

```{r}
library(RColorBrewer)
cols<-brewer.pal(3, "BuGn")
cols
```



### ggplot2

Implementation of the Grammar of Graphics


#### qplot() function

Looks for data in a data frame.

```{r}
library(ggplot2)
str(mpg)
```


```{r}
qplot(displ, hwy, data=mpg)
```

```{r}
qplot(displ, hwy, data=mpg, color = drv)
```



```{r}
qplot(displ, hwy, data=mpg, geom = c("point", "smooth"))
```

We can also get a histogram with qplot

```{r}
qplot(hwy, data = mpg, fill = drv)
```


#### Facets

Facets are like panels in lattice

```{r}
qplot(displ, hwy, data = mpg, facets = . ~ drv)
```

```{r}
qplot(hwy, data = mpg, facets = drv ~ ., binwidth = 2)
```

### ggplot2 part 2

Basic components of ggplot2 plot

- data frame
- aesthetic mappings
- geoms
- facets
- stats
- scales
- coordinate system


#### Building up in layers:

`head(maacs[, 1:3])
g <- ggplot(maacs, aes (logpm25, NocturnalSympt))
summary(g)
g + geom_point()
g + geom_point() + geom_smooth()
g + geom_point() + geom_smooth(method = "lm")
g + geom_point() + facet_grid (. ~bmicat) + geom_smooth(method = "lm")
`
##### Annotation

Labels:

- xlab()
- ylab()
- labs()
- ggtitle()

global theme:

theme(legend.position = "none")

Two standard appearence themes are included:

- theme_gray(): default one with gray background.
- theme_bw()

`
g + geom_point(aes(color = bmicat), size=4, alpha = 1/2))
`

aes is aesthetics.

### ggplot2 part 5

```{r}
library(ggplot2)
testdat <- data.frame(x = 1:100, y = rnorm(100))
testdat[50,2] <- 100 ## outlier
plot(testdat$x, testdat$y, type = "l", ylim = c(-3,3))
```



```{r}
g <- ggplot(testdat, aes(x = x, y = y))
g + geom_line()
```

```{r}
g + geom_line() + coord_cartesian(ylim = c(-3,3))
```


#### Quantifying continuos variables

Calculate the deciles of the data

`cutpoints <- quantile(maacs$logno2_new, seq(0,1, length = 4), na.rm = TRUE)`

Cut the data at the deciles and create a new factor variable

`maacs$no2dec <- cut(maac$logno2_new, cutpoints)`

See the levels of the newly created factor variable

`levels(maacs$no2dec)`

Final code of example:

```
g <- ggplot (maacs, aes(logpm24, NocturnalSympt))

## Add layers

g + geom_point(alpha = 1/3)
  + facet_wrap(bmicat ~no2dec, nrow = 2, ncol = 4)
  + geom_smooth(method = "lm", se=FALSE, col="steelblue")
  + theme_bw(base_family = "Avenir", base_size = 10)
  + labs(x = expression("log " * PM[2.5])
  + labs(y = "Nocturnal Symptoms")
  + labs(title = "MAACS Cohort")
```

#### ggplot2 exercises (I)

Using the mpg data provided by ggplot2 itself (I think...)

```{r}
qplot(displ, hwy, data=mpg, color=drv, geom=c("point", "smooth"))
```

```{r}
qplot(y=hwy, data = mpg, color = drv)

```

We can also do some boxplotting
```{r}
qplot(drv, hwy, data = mpg, geom = "boxplot")
```

and group even for color / manufacturer

```{r}
qplot(drv, hwy, data = mpg, geom = "boxplot", color=manufacturer)
```


#### We can draw histograms

```{r}
qplot(hwy, data=mpg, fill=drv)
```

#### Facets or panels

```{r}
qplot(displ, hwy, data=mpg, facets= . ~drv) ## left of ~ is rows, right of ~ is columns
```

Now with histograms

```{r}
qplot(hwy, data=mpg, facets= drv ~ ., binwidth = 2)
```

#### ggplot2 exercises (II)

A three facet plot, for each drive type.

```{r}
qplot(displ, hwy, data = mpg, geom=c("point", "smooth"), facets = .~drv)
```


We can build something similar step by step... let's see:

```{r}
g <- ggplot(mpg, aes(displ, hwy))
g+geom_point()
```

geom_point adds a layer.

```{r}
g+geom_point()+geom_smooth()
```


we can generate a regression model instead of confidence interval:

```{r}
g+geom_point()+geom_smooth(method="lm")
```

Now with facets

```{r}
g+geom_point()+geom_smooth(method="lm")+facet_grid(. ~ drv)
```

##### Modifying aesthetics

```{r}
g+geom_point(color="pink", size=4, alpha=1/2)
```

```{r}
g+geom_point(size=4, alpha=1/2, aes(color=drv))
```


```{r}
g+geom_point(aes(color=drv))+labs(title="Swirl Rules!")+labs(x="Displacement", y="Hwy Mileage")
```


```{r}
g+geom_point(aes(color=drv), size=2, alpha=1/2)+geom_smooth(size=4, linetype=3, method="lm", se=FALSE)
```

```{r}
g+geom_point(aes(color=drv))+theme_bw(base_family="Times")
```

#### More complicated plots

```{r}
g <- ggplot(data=mpg, aes(x=displ, y=hwy, color=factor(year)))
g+geom_point()+facet_grid(drv~cyl, margins=TRUE)
```

```{r}
g+geom_point()+facet_grid(drv~cyl, margins=TRUE)+geom_smooth(method="lm", se=FALSE, size=2, color="black")+labs(x="Displacement", y="Highway Mileage", title="Swirl Rules!")

```



